Clone and graft:\\ Testing scientific applications as they are built

### 2015-01-23 turcksin-spring-mechanics/paper.tex
add two points

Developing testing schemes for libraries is reasonably well understood. So,
when we ventured in the area of building applications for scientific purposes
on top of the libraries we had created, we had initially thought that one can
use the same approaches as one uses for libraries -- but this turned out to
not work.

Libraries are relatively easy to test. They export large numbers of
functions; one can write a simple \texttt{main()} function that sets up a
couple of data structures, calls one of these functions individually, and
verifies the correctness of the output; do
this for every function or class in the library and you've got a good
testsuite. Such tests are
often called \textit{unit tests} and are typically complemented by integration
tests to verify that combinations of classes work well together. Over the
years we have written about 3,000 of these that are run in multiple
configurations with every single commit to \dealii{} -- on multiple machines,
with different compilers and dependencies, and several times a day. Despite
all of the experience we have gained building this machinery around \dealii{},
the same strategy does not work for whole applications because applications do
not usually export many possible entry points. Rather, they usually only
present a single interface -- the command line, a graphical user interface, or
input files -- and then run through a large fraction of the entire code base
in computing output. One cannot typically just call a single function in an
application by presenting a magic input file, in the same way as one would
call a function in a library by presenting a magic \texttt{main()} function in
a unit test.

@@

Let's implement a program that can solve this problem. Because we want it
to be simple and accessible, we're going to use Python as it's widely used and
because it comes with a number of tools that are going to make our life
simpler by keeping the program short.%
\footnote{In order to run this program, you will have to have the Python
  NumPy, SciPy, and matplotlib packages installed on your system, along with
  the basic Python interpreter.}
The program we'll show you in a second reads its input parameters from a file in JSON format%
\footnote{While certainly more sane compared to XML, JSON is not the prettiest format to write input files in and we
  don't recommend you use it for your programs (it is typically intended for
  exchange of data between programs, not between humans and computers). But it
  will serve for our purposes here primarily because Python has a built-in
  parser for it that makes our program pleasantly free of the tedium of
  parsing input files.} (see
\url{http://en.wikipedia.org/wiki/JSON}).
Here is an example that you can find in the file \texttt{tests/testcase-1.json}:
\lstinputlisting[frame=single,numbers=left,basicstyle=\footnotesize]{tests/testcase-1.json}
In JSON, the name of a parameter is to the left of the colon and its value is
to the right; name/value pairs are separated by commas. The value of a
parameter can be a plain number, or it can be an array if enclosed in
brackets. Arrays can also be nested, which is what we use here for the initial
positions and velocities of our two bodies, each of which lives in
three-dimensional space. With this explanation, it is clear that the input file
specifies properties of the two masses as $\mathbf x_{1,0}=(1,2,3)^T, \mathbf
x_{2,0}=(4,5,6)^T, \mathbf v_{1,0}=(0,0,0)^T, \mathbf v_{2,0}=(1,1,1)^T,
m_1=13.5, m_2=29.75$ and $D=42, L=2.25$ for the spring that connects them, 
thereby completely determining everything in equations
\eqref{eq:ode}--\eqref{eq:ic2}.

>>>>

Developing testing schemes for libraries is reasonably well understood. So,
when we ventured in the area of building \textit{applications} for scientific purposes
on top of the libraries we had created, we had initially thought that one can
use the same approaches as one uses for libraries -- but this turned out to
not work.

Libraries are relatively easy to test. They export large numbers of
functions; one can write a simple \texttt{main()} function that sets up a
couple of data structures, calls one of these functions individually, and
verifies the correctness of the output; do
this for every function or class in the library and you've got a good
testsuite. Such tests are
often called \textit{unit tests} and are typically complemented by integration
tests to verify that combinations of classes work well together. Over the
years we have written about 3,000 of these that are run in multiple
configurations with every single commit to \dealii{} -- on multiple machines,
with different compilers and dependencies, and several times a day. Despite
all of the experience we have gained building this machinery around \dealii{},
the same strategy does not work for whole applications because applications do
not usually export many possible entry points. Rather, they usually only
present a single interface -- the command line, a graphical user interface, or
input files -- and then run through a large fraction of the entire code base
in computing output. One cannot typically just call a single function in an
application by presenting a magic input file, in the same way as one would
call a function in a library by presenting a magic \texttt{main()} function in
a unit test. There are of course more advanced ways of testing features of an
application in isolation by replacing other parts of the code during testing by so-called
mock or stub objects. We will not discuss this here though.

@@

Let's implement a program that can solve this problem. Because we want it
to be simple and accessible, we're going to use Python as it's widely used and
because it comes with a number of tools that are going to make our life
simpler by keeping the program short.%
\footnote{In order to run this program, you will have to have the Python
  NumPy, SciPy, and matplotlib packages installed on your system, along with
  the basic Python interpreter.}
Note that testing is even more important in dynamically typed languages like
Python, because you do not have the compiler to help you find bugs at compile 
time.

The program we'll show you in a second reads its input parameters from a file in JSON format%
\footnote{While certainly more sane compared to XML, JSON is not the prettiest format to write input files in and we
  don't recommend you use it for your programs (it is typically intended for
  exchange of data between programs, not between humans and computers). But it
  will serve for our purposes here primarily because Python has a built-in
  parser for it that makes our program pleasantly free of the tedium of
  parsing input files.} (see
\url{http://en.wikipedia.org/wiki/JSON}).
Here is an example that you can find in the file \texttt{tests/testcase-1.json}:
\lstinputlisting[frame=single,numbers=left,basicstyle=\footnotesize]{tests/testcase-1.json}
In JSON, the name of a parameter is to the left of the colon and its value is
to the right; name/value pairs are separated by commas. The value of a
parameter can be a plain number, or it can be an array if enclosed in
brackets. Arrays can also be nested, which is what we use here for the initial
positions and velocities of our two bodies, each of which lives in
three-dimensional space. With this explanation, it is clear that the input file
specifies properties of the two masses as $\mathbf x_{1,0}=(1,2,3)^T, \mathbf
x_{2,0}=(4,5,6)^T, \mathbf v_{1,0}=(0,0,0)^T, \mathbf v_{2,0}=(1,1,1)^T,
m_1=13.5, m_2=29.75$ and $D=42, L=2.25$ for the spring that connects them, 
thereby completely determining everything in equations
\eqref{eq:ode}--\eqref{eq:ic2}.
